On the Cram\'{e}r-Rao bound for polynomial phase signals

### 2012-10-17 paper.tex
minor

A uniformly sampled, constant amplitude, polynomial-phase signal of order $m$ has the model~\cite{Peleg_DPT_1995, Peleg1991_CRB_PPS_1991, Peleg1991_est_class_PPS_1991, Rihaczek1996},
\[
s_n(b) = \alpha \exp\left( j \sum_{k = 0}^{m}{b_k (\Delta n)^k}\right)
\]
where $n \in \ints$, $\alpha$ is the signal amplitude (a positive real number), $b=(b_0, \dots, b_m)$ is a vector containing the polynomial coefficients, $j = \sqrt{-1}$ and $\Delta$ is the interval between consecutive samples.  We will assume, without loss of generality, that $\Delta = 1$.

@@

We make the assumption that the noise sequence $\{ w_n \}$ is independent and identically distributed complex Gaussian, each element having independent real and imaginary parts with zero mean and variance $\sigma^2$.  The probability density function of $w_n$ is 
\[
p(x) = \frac{1}{4 \pi \sigma^2} \exp\left( -\frac{\abs{x}^2}{2\sigma^2} \right)
\]
where $\abs{x}^2$ is the squared magnitude of the complex number $x$.  We wish to estimate $\beta$ from $y_1, \dots, y_N$.  The likelihood function is
\[
L(b) =  \frac{1}{(4 \pi \sigma^2)^N} \exp\left( - \frac{1}{2\sigma^{2}} \sum_{n=1}^N \abs{y_n -  s_n(b) }^2 \right).
\]
%The maximum likelihood estimators of $\alpha, \beta_0, \dots, \beta_m$ are the maximisers of $L$.
% , that is
% \[
% (\hat{\alpha}, \hat{\beta}_0, \dots, \hat{\beta}_m) = \arg\max_{a \in \reals, (b_0, \dots, b_m) \in B} L(a, b_0, \dots, b_m) 
% \]  
% where $B \subset \reals^{m+1}$ describes a region of polynomial phase coefficients for which the estimation problem is identifiable~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase,McKilliam_LSU_polyest_part1_2012}.  We assume that $(\beta_0, \dots, \beta_m) \in B$.
For $i,k \in \{0, \dots, m\}$ define functions 
\[
f_{ik}(b) = f_{ki}(b) = -\expect \frac{ \partial^2 \log L(b) }{ \partial b_{i} \partial b_{k}}.
\]  
The Fisher information matrix $F$ is the $m+1$ by $m+1$ matrix with elements $F_{i k} = f_{ik}(\beta)$.  The elements are
\[
F_{ik} = F_{ki} = \frac{\alpha^2}{\sigma^2}\sum_{n=1}^{N} n^{i}n^{k}.
\]

@@

For a polynomial $g(x) = b_0 + b_1 x + \dots + b_m x^m$ of order at most $m$ denote by $\coef(g) = (b_0, b_1, \dots, b_m)$ the vector of length $m+1$ containing its coefficients.  Let $p_0, \dots, p_m$ be a family of linearly independent polynomials.  By linearly independent it is meant that any polynomial of order at most $m$ can be written as a linear combination $c_0p_0 + c_1 p_1 + \dots + c_m p_m$ where $c_1, \dots, c_m$ are real numbers.  Observe that $p_0, \dots, p_m$ are linearly independent if and only if the vectors $\coef(p_0), \dots, \coef(p_m)$ are linearly independent.  We call $p_0, \dots, p_m$ a \emph{basis} for the space of polynomials of order $m$.  We call the polynomials $n^0, n^1, \dots, n^m$ the \emph{standard basis}.

@@

Put,
\[
r_n(c) = \alpha \exp\left( j \sum_{k = 0}^{m}{c_k p_k(n)}\right)
\]
so that $r_n(c)$ is a polynomial phase signal expressed in terms of the basis $p_0, \dots, p_m$.  Correspondingly $s_n(b)$ is a polynomial phase expressed in terms of the standard basis $n^0, n^1, \dots, n^m$.  The two signals are related by
\[
r_n(c) = s_n(cP) = s_n(b).
\]
The observed samples $y_1,\dots,y_N$ can be written using $r_n$ as
\[
y_n = r_n(\gamma) + w_n = s_n(\gamma P) + w_n
\]
where $\beta = \gamma P$.  We can now consider the problem of estimating $\gamma$ instead of $\beta$.  If $\hat{\beta}$ is an estimator of $\beta$ then the corresponding estimator of $\gamma$ is $\hat{\gamma} = \hat{\beta}P^{-1}$.  If $C$ is the covariance of $\hat{\beta}$ then the covariance of $\hat{\gamma}$ is $(P^{-1})^\prime C P^{-1}$ where $^\prime$ is the transpose.

@@

Observe that the polynomials $q_0,q_1,\dots$ satisfy the recursion
\[
q_s(x) = \frac{x-s}{s} q_{s-1}(x), \qquad q_{0}(x) = 1.
\]
If $q_s(x) = q_{s0} + q_{s1}x + \dots, q_{ss}x^s$ so that $q_{s0},q_{s1},\dots,q_{ss}$ are the coefficients of $q_{s}$, then
\[
q_{si} = \frac{1}{s}q_{s-1,i-1} - q_{s-1,i}
\]
where $q_{00}=1$ and $q_{0i}=0$ for $i \neq 0$ and $q_{si} = 0$ if either of $s$ or $i$ is negative.

>>>>

A uniformly sampled, constant amplitude, polynomial-phase signal of order $m$ has the model~\cite{Peleg_DPT_1995, Peleg1991_CRB_PPS_1991, Peleg1991_est_class_PPS_1991, Rihaczek1996},
\[
s_n(b) = \alpha \exp\left( j \sum_{k = 0}^{m}{b_k (\Delta n)^k}\right)
\]
where $n \in \ints$, $\alpha$ is the signal amplitude (a positive real number), $b=(b_0, \dots, b_m)$ is a vector containing the polynomial coefficients (each a real number),  and $\Delta$ is the interval between consecutive samples.  We will assume, without loss of generality, that $\Delta = 1$.

@@

We make the assumption that the noise sequence $\{ w_n \}$ is independent and identically distributed complex Gaussian, each element having independent real and imaginary parts with zero mean and variance $\sigma^2$.  The probability density function of $w_n$ is 
\[
p(x) = \frac{1}{2 \pi \sigma^2} \exp\left( -\frac{\abs{x}^2}{2\sigma^2} \right)
\]
where $\abs{x}^2$ is the squared magnitude of the complex number $x$.  We wish to estimate $\beta$ from $y_1, \dots, y_N$.  The likelihood function is
\[
L(b) =  \frac{1}{(4 \pi \sigma^2)^N} \exp\left( - \frac{1}{2\sigma^{2}} \sum_{n=1}^N \abs{y_n -  s_n(b) }^2 \right).
\]
%The maximum likelihood estimators of $\alpha, \beta_0, \dots, \beta_m$ are the maximisers of $L$.
% , that is
% \[
% (\hat{\alpha}, \hat{\beta}_0, \dots, \hat{\beta}_m) = \arg\max_{a \in \reals, (b_0, \dots, b_m) \in B} L(a, b_0, \dots, b_m) 
% \]  
% where $B \subset \reals^{m+1}$ describes a region of polynomial phase coefficients for which the estimation problem is identifiable~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase,McKilliam_LSU_polyest_part1_2012}.  We assume that $(\beta_0, \dots, \beta_m) \in B$.
For $i,k \in \{0, \dots, m\}$ define functions 
\[
f_{ik}(b) = f_{ki}(b) = -\expect \frac{ \partial^2 \log L(b) }{ \partial b_{i} \partial b_{k}}.
\]  
The Fisher information matrix $F$ is the $m+1$ by $m+1$ matrix with elements $F_{i k} = f_{ik}(\beta)$.  The elements are
\[
F_{ik} = F_{ki} = \frac{\alpha^2}{\sigma^2}\sum_{n=1}^{N} n^{i}n^{k}.
\]

@@

For a polynomial $g(x) = b_0 + b_1 x + \dots + b_m x^m$ of order at most $m$ denote by $\coef(g) = (b_0, b_1, \dots, b_m)$ the vector of length $m+1$ containing its coefficients.  Let $p_0,\dots, p_m$ be a family of linearly independent polynomials of orders at most $m$.  By linearly independent it is meant that any polynomial of order at most $m$ can be written as a linear combination $c_0p_0 + c_1 p_1 + \dots + c_m p_m$ where $c_1, \dots, c_m$ are real numbers.  Observe that $p_0, \dots, p_m$ are linearly independent if and only if the vectors $\coef(p_0), \dots, \coef(p_m)$ are linearly independent.  We call $p_0, \dots, p_m$ a \emph{basis} for the space of polynomials of order $m$.  We call the polynomials $x^0, x^1, \dots, x^m$ the \emph{standard basis}.

@@

Put,
\[
r_n(c) = \alpha \exp\left( j \sum_{k = 0}^{m}{c_k p_k(n)}\right)
\]
so that $r_n(c)$ is a polynomial phase signal expressed in terms of the basis $p_0, \dots, p_m$.  Correspondingly $s_n(b)$ is a polynomial phase signal expressed in terms of the standard basis $x^0, x^1, \dots, x^m$.  The two signals are related by
\[
r_n(c) = s_n(cP) = s_n(b).
\]
The observed samples $y_1,\dots,y_N$ can be written using $r_n$ as
\[
y_n = r_n(\gamma) + w_n = s_n(\gamma P) + w_n
\]
where $\beta = \gamma P$.  We can now consider the problem of estimating $\gamma$ instead of $\beta$.  If $\hat{\beta}$ is an estimator of $\beta$ then the corresponding estimator of $\gamma$ is $\hat{\gamma} = \hat{\beta}P^{-1}$.  If $C$ is the covariance of $\hat{\beta}$ then the covariance of $\hat{\gamma}$ is $(P^{-1})^\prime C P^{-1}$ where $^\prime$ is the transpose.

@@

Observe that the polynomials $q_0,q_1,\dots$ satisfy the recursion
\[
q_s(x) = \frac{x-s}{s} q_{s-1}(x), \qquad q_{0}(x) = 1.
\]
If $q_s(x) = q_{s0} + q_{s1}x + \dots, q_{ss}x^s$ so that $q_{s0},q_{s1},\dots,q_{ss}$ are the coefficients of $q_{s}$, then
\[
q_{si} = \frac{1}{s}q_{s-1,i-1} - q_{s-1,i}
\]
where $q_{00}=1$ and $q_{0i}=0$ for $i \neq 0$ and $q_{si} = 0$ if any of $s$ or $i$ is negative.
